Transition path sampling based calculation of free energies of enzymatic reactions

### 2022-04-12 tex/manuscript.tex
added some more figures and some text

TPS was developed as a reaction-coordinate free, statistical method 
which samples trajectories that connect the stable reactant and product states 
guaranteeing the occurence of the rare barrier crossing events. These trajectories 
are sampled using a Monte carlo procedure leading to an ensemble of transition
pathways that are real dynamical trajectories which contain information regarding the 
mechanism and kinetics of the reaction.

TPS in particular is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
a complicated entity in large biological systems. TPS has been successfully 
applied to study problems such as enzyme catalysis, protein folding, etc. 
 In this Monte-Carlo procedure in the trajecotry space
rejected moves correspond to non-reactive trajectories. But these trajectories are 
required to obtain the equilibrium distribution function.

The TPS ensemble consists of reactive trajectories which connect a reactant state to
a product state by going through the transition state and the probabilities of configurations thus sampled 
do not replicate the equilibrium probabilities. To calculate equilibrium properties such as the free energy 
of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
required which is not obtained with straightforward TPS. \cite{Dellago09AdvCompSimAppp167}

@@

%------------------------------------------------------------------------------ We performed computational studies on two enzymes where we have shown
that fast dynamics is part of the reaction mechanism and calculated the electric field near
the bond-breaking event.

@@

%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
%------------------------------------------------------------------------------ 
Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
reaction between Adenosine triphosphaste and L-methionine which leads to the 
formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
The mechanism of this reaction is determined to follow a SN$_2$ transition state
with the sulphur atom of the L-methionine as the nucleophile with the 
triphosphate leaving group.

@@

\begin{figure}
\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}
%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.22]{./figures/ada-protein.png}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.9]{./figures/ada-fenergy.pdf}
\end{figure}

%\begin{figure}[ht]
%\centering
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/reactant-state.png}
%\caption{Reactant state}
%\label{fig:minipage1}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/int1.png}
%\caption{Intermediate state}
%\label{fig:minipage2}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/product-state.png}
%\caption{Product state}
%\label{fig:minipage1}
%\end{minipage}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

>>>>

TPS in particular is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
complex for large biological systems.
TPS can be considered as a generalization of the Monte Carlo sampling of the 
configuration space to the space of trajectories where the collected ensemble of 
reactive pathways are real dynamical trajectories compared to articial trajectories
generated using bias forces. The TPS ensemble can be used to obtain kinetic information
such as the rate of the reaction as well as the transition state, reaction
coordinate and the reaction mechanism. TPS has been successfully 
applied to study problems such as enzyme catalysis, protein folding, etc.

Since TPS focuses on collecting only reactive trajectories the information contained 
in the non-reactive trajectories which is essential for describing the 
equilibrium distribution is missing. 
Methods that can be used to simultaneously calculate
the dynamical as well as equilibrium properties of a system within 
a certain set of asssumptions have garnered interest in recent times. \cite{Suarez14JChemTheoryComput10p2658}
Transition interface sampling (TIS) is a variant of TPS which can be used for 
this purpose. \cite{vanErp03JChemPhys118p7762}

A common theme among such methods is to divide the reaction coordinate into
windows within which independent sampling can be performed and interpolated
in a way that can used to calculate equilibrium properties

To calculate equilibrium properties such as the free energy 
of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
required which is not obtained with straightforward TPS. \cite{Dellago09AdvCompSimAppp167}

@@

@@

%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
%------------------------------------------------------------------------------ 
Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
reaction between adenosine triphosphaste and L-methionine which leads to the 
formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
The mechanism of this reaction is determined to follow a SN$_2$ transition state
with the sulphur atom of the L-methionine as the nucleophile with the 
triphosphate leaving group.

@@

\begin{figure}
\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}
We used transition path sampling to generate and
analyze the ensemble of reactive trajectories. As order
parameters that define the reactant and product basins, we
used the donor-hydride and acceptor-hydride distances. The
reactant basin was defined as a donor-hydride distance smaller
than 1.6  and the product basin as an acceptor-hydride
distance smaller than 1.6 . To seed the generation of the path
ensemble, we generated a biased trajectory by applying a
harmonic force between the hydride and the acceptor and
allowed the system to propagate for 250 fs. New trajectories
were generated by the shooting algorithm: we perturbed the
momenta of a randomly chosen slice of a reactive trajectory,
and using the new momenta, we allowed the system to
propagate forward and backward in time for 250 fs. A total of
120 reactive trajectories were generated.
%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.22]{./figures/ada-protein.png}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.6]{./figures/ada-fenergy.pdf}
\end{figure}

%\begin{figure}[ht]
%\centering
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/reactant-state.png}
%\caption{Reactant state}
%\label{fig:minipage1}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/int1.png}
%\caption{Intermediate state}
%\label{fig:minipage2}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/product-state.png}
%\caption{Product state}
%\label{fig:minipage1}
%\end{minipage}
%\end{figure}
%------------------------------------------------------------------------------ 
\section{Conclusions}
%------------------------------------------------------------------------------ 
We have discussed the application of a scheme to calculate the 
free energies of enzymatic reactions within the TPS method using the MAT2A
and the ADA enzymes as examples.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
